---
title: "Beerdata"
author: "Ndede, George M."
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Load Libraries
library(dplyr)      # for string functions//Heavy use of this library
library(ggplot2)    # for plots
library(kableExtra) # for table formatting// library to make the document more presentable
library(knitr)      # for presenting in html
library(stringr)    # for str_trim function
library(ggthemes)  # for themes in ggplot
library(tidyr)
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function

```

###A. Clean, prepare and load the data
```{r echo = TRUE}

####Que1. First read the two data sets into data frames beer Data Frame and brewery Data Frame from csv files. 
### The columns (variables) of each data frame were custom named to be more readable and to provide a handle to merge the data.Create column for Brewery ID that is common to both datasets.
# Read beer data from csv file and put into brewery Data Frame data frame.Then rename columns to be more readable.

BeerDataFrame <- read.csv("./DataFilesFolder/Beers.csv",header=TRUE,sep=",",stringsAsFactors = TRUE, encoding = "UTF-8")
```

## Renaming Brewery_id to Brew_ID to satisfy merging requirement
```{r echo = TRUE}
names(BeerDataFrame) <- c("Beer_Name","Beer_ID","Alcohol_By_Volume","Bitterness_Units","Brewery_ID",
                   "Beer_Style","Ounces") # rename table columns
#BeerDataFrame
```

# Taking a quick look at the dataset
```{r echo = TRUE}
glimpse(BeerDataFrame)
```

# Read Brewery data from csv file then put into Brewery Data Frame. Then rename columns to be more readable.
```{r echo = TRUE}
BreweryDataFrame <- read.csv("./DataFilesFolder/Breweries.csv",header=TRUE,sep=",",stringsAsFactors = FALSE)

# Renaming Brewery_id to Brew_ID to satisfy merging requirement
names(BreweryDataFrame) <- c("Brewery_ID","Brewery_Name","City","State") # rename table columns


# Taking a quick look at the data set summary in display
glimpse(BreweryDataFrame)
```


#1: How many breweries are present in each state?
```{r echo = TRUE}
# How many breweries are present in each state?
BreweryperState <- table(BreweryDataFrame$State)
BreweryperState
```


#2:
#Merged Brewery and Beer data into the BrewDataFrame
```{r echo = TRUE}
AllBeerDataFrame <- merge.data.frame(BreweryDataFrame, BeerDataFrame, by="Brewery_ID")
#AllBeerDataFrame
```

#ii): # print first 6 rows of AllBeerDataFrame
```{r echo = TRUE}
head(AllBeerDataFrame, n=6)
```

#iii) # print last 6 rows of AllBeerDataFrame
```{r echo = TRUE}
tail(AllBeerDataFrame, n=6)
```



#3. Address the missing values in each column.
# Code to replace missing values with the mean // 
```{r echo = TRUE}
BeerDataFrame_impute_mean <- data.frame(
  sapply(
    BeerDataFrame,
    function(x) ifelse(is.na(x),
                       mean(x, na.rm = TRUE),
                       x)))
#Output of the cleaed dataset without missing values
#BeerDataFrame_impute_mean
```



